\Large\textbf{General Stage Untangling with Applications to Algorithm Derivation}

### 2014-02-21 paper/paper.tex
Added a better semantics.  Refactored some figures using fancy latex macros.
Refactored splitting rules.  Fixed up some proofs.

\begin{figure}
\caption{\corelang~Static Semantics}
\label{fig:coreStatics}
\begin{mathpar}
\infertypesone [\rmunit] 			{\typesone {()}{\rm unit}}{\cdot} \and
\infertypesone [\mathrm{int}]		{\typesone {i}{\rm int}}{\cdot} \and
\infertypesone [\mathrm{bool}]	{\typesone {b} {\rm bool}}{\cdot} \and
\infertypesone [\mathrm{hyp}]		{\typesone x A}{x : A^\bbone \in \Gamma} \and
\infertypestwo [\rmunit]			{\typestwo {()}{\rm unit}}{\cdot} \and
\infertypestwo [\mathrm{int}]		{\typestwo {i}{\rm int}}{\cdot} \and
\infertypestwo [\mathrm{bool}]	{\typestwo {b} {\rm bool}}{\cdot} \and
\infertypestwo [\mathrm{hyp}]		{\typestwo x A}{x : A^\bbtwo \in \Gamma} \and
\infertypesone [\mathrm{let}]		{\typesone{\letin{x}{e_1}{e_2}}{B}}{\typesone{e_1} A & \typesone [\Gamma,x:A^\bbone] {e_2} B} \and
\infertypestwo [\mathrm{let}]		{\typestwo{\letin{x}{e_1}{e_2}}{B}}{\typestwo{e_1} A & \typestwo [\Gamma,x:A^\bbtwo] {e_2} B} \and
\infertypesone [\times\mathrm{I}]		{\typesone {(e_1,e_2)} {A\times B}}{\typesone {e_1} A & \typesone {e_2} B} \and
\infertypesone [\times\mathrm{E_1}]	{\typesone{\pi_1~e} A}{\typesone e {A\times B}} \and
\infertypesone [\times\mathrm{E_2}]	{\typesone{\pi_2~e} B}{\typesone e {A\times B}} \and
\infertypestwo [\times\mathrm{I}]		{\typestwo {(e_1,e_2)} {A\times B}}{\typestwo {e_1} A & \typestwo {e_2} B} \and
\infertypestwo [\times\mathrm{E_1}]	{\typestwo{\pi_1~e} A}{\typestwo e {A\times B}} \and
\infertypestwo [\times\mathrm{E_2}]	{\typestwo{\pi_2~e} B}{\typestwo e {A\times B}} \and
\infertypesone [\mathrm{if}] {\typesone{\ifthen{e_1}{e_2}{e_3}} A}{\typesone{e_1} {\rm bool} & \typesone {e_2} A & \typesone {e_3} A} \and
\infertypestwo [\mathrm{if}] {\typestwo{\ifthen{e_1}{e_2}{e_3}} A}{\typestwo{e_1} {\rm bool} & \typestwo {e_2} A & \typestwo {e_3} A} \and
\infertypesone [\fut\mathrm{I}]	{\typesone{\next~e}{\fut A}}{\typestwo e A} \and
\infertypestwo [\fut\mathrm{E}]	{\typestwo {\prev~e} A}{\typesone e {\fut A}} \and
\end{mathpar}
\end{figure}

@@

\begin{figure}
\caption{\corelang~Erasure Dynamic Semantics}
\label{fig:erasureSemantics}
\begin{mathpar}
\inferreduceone [\rmunit]				{\reduceone {()}{()}}{\cdot} \and
\inferreduceone [\mathrm{int}]		{\reduceone{i}{i}}{\cdot} \and
\inferreduceone [\mathrm{bool}]		{\reduceone{b}{b}}{\cdot} \and
\inferreduceone [\mathrm{let}]			{\reduceone{\letin{x}{e_1}{e_2}}{v_2}}{\reduceonesub [1] & \reduceone{[v_1/x]e_2}{v_2}} \and
\inferreducetwo [\rmunit]				{\reducetwo {()}{()}}{\cdot} \and
\inferreducetwo [\mathrm{int}]		{\reducetwo{i}{i}}{\cdot} \and
\inferreducetwo [\mathrm{bool}]		{\reducetwo{b}{b}}{\cdot} \and
\inferreducetwo [\mathrm{let}]			{\reducetwo{\letin{x}{e_1}{e_2}}{v_2}}{\reducetwosub [1] & \reducetwo{[v_1/x]e_2}{v_2}} \and
\inferreduceone [\times\mathrm{I}]	{\reduceone{(e_1,e_2)}{\valprod{v_1}{v_2}}}{\reduceonesub [1] & \reduceonesub [2]} \and
\inferreduceone [\times\mathrm{E_1}]	{\reduceone{\pi_1~e}{v_1}}{\reduceone{e}{\valprod{v_1}{v_2}}} \and
\inferreduceone [\times\mathrm{E_2}]	{\reduceone{\pi_2~e}{v_2}}{\reduceone{e}{\valprod{v_1}{v_2}} } \and
\inferreducetwo [\times\mathrm{I}]	{\reducetwo{(e_1,e_2)}{\valprod{v_1}{v_2}}}{\reduceonesub [1] & \reduceonesub [2]} \and
\inferreducetwo [\times\mathrm{E_1}]	{\reducetwo{\pi_1~e}{v_1}}{\reduceone{e}{\valprod{v_1}{v_2}}} \and
\inferreducetwo [\times\mathrm{E_2}]	{\reducetwo{\pi_2~e}{v_2}}{\reduceone{e}{\valprod{v_1}{v_2}}} \and
\inferreduceone [\mathrm{if_1}] 		{\reduceone{\ifthen{e_1}{e_2}{e_3}}{v}}{\reduceone{e_1}{\bf true} & \reduceone{e_2}{v}} \and
\inferreduceone [\mathrm{if_2}] 		{\reduceone{\ifthen{e_1}{e_2}{e_3}}{v}}{\reduceone{e_1}{\bf false} & \reduceone{e_3}{v}} \\
\inferreducetwo [\mathrm{if_1}] 		{\reducetwo{\ifthen{e_1}{e_2}{e_3}}{v}}{\reducetwo{e_1}{\bf true} & \reducetwo{e_2}{v}} \and
\inferreducetwo [\mathrm{if_2}] 		{\reducetwo{\ifthen{e_1}{e_2}{e_3}}{v}}{\reducetwo{e_1}{\bf false} & \reducetwo{e_3}{v}} \\
\inferreduceone [\fut\mathrm{I}]		{\reduceone{\next~e}{\valnext~v}}{\reducetwosub} \and
\inferreducetwo [\fut\mathrm{E}]		{\reducetwo{\prev~e} v}{\reduceone e {\valnext~v}} 
\end{mathpar}
\end{figure}

%\subsection{Better Dynamic Semantics}
%
%Introduce the dynamics semantics.  Figure \ref{fig:coreDynamics}.  It's small step.  We figured this out on our own, but it really should exist in the partial evaluation literature somewhere.
%
%\newcommand {\stepo} {\hookrightarrow_\bbone} 
%\newcommand {\stept} {\hookrightarrow_\bbtwo} 
%\newcommand {\pco} {~\mathrm{pc}\bbone} 
%\newcommand {\fco} {~\mathrm{fc}\bbone} 
%\newcommand {\pct} {~\mathrm{pc}\bbtwo} 
%\newcommand {\fct} {~\mathrm{fc}\bbtwo} 
%
%\begin{figure}
%\caption{\corelang~Partially Computed Values (pc$\bbone$ and pc$\bbtwo$)}
%\label{fig:coreDynamics}
%\begin{mathpar}
%\infer {() \pco}{\cdot} \and
%\infer {i \pco}{\cdot} \and
%\infer {b \pco}{\cdot} \and
%\infer {(v_1, v_2) \pco}{v_1 \pco & v_2 \pco} \and
%\infer {\next~e \pco}{e \pct} \\
%\infer {() \pct}{\cdot} \and
%\infer {i \pct}{\cdot} \and
%\infer {b \pct}{\cdot} \and
%\infer {(v_1, v_2) \pct}{v_1 \pct & v_2 \pct} \and
%\infer {\pi_i e \pct}{e \pct} \and
%\infer {x \pct}{\cdot} \and
%\infer {\letin{x}{e_1}{e_2} \pct}{e_1 \pct & e_2 \pct} \and
%\infer {\ifthen{e_1}{e_2}{e_3} \pct}{e_1 \pct & e_2 \pct & e_3 \pct} \and
%\infer {\prev~e \pct}{e \pco} 
%\end{mathpar}
%\end{figure}
%
%\begin{figure}
%\caption{\corelang~Fully Computed Values (fc$\bbone$ and fc$\bbtwo$)}
%\label{fig:coreDynamics}
%\begin{mathpar}
%\infer {() \fco}{\cdot} \and
%\infer {i \fco}{\cdot} \and
%\infer {b \fco}{\cdot} \and
%\infer {(v_1, v_2) \fco}{v_1 \fco & v_2 \fco} \and
%\infer {\next~e \fco}{e \fct} \\
%\infer {() \fct}{\cdot} \and
%\infer {i \fct}{\cdot} \and
%\infer {b \fct}{\cdot} \and
%\infer {(v_1, v_2) \fct}{v_1 \fct & v_2 \fct} 
%\end{mathpar}
%\end{figure}
%
%\begin{figure}
%\caption{\corelang~Transitions}
%\label{fig:coreDynamics}
%\begin{mathpar}
%\infer {\ifthen{e_1}{e_2}{e_3} \stepo \ifthen{e_1'}{e_2}{e_3}}{e_1 \stepo e_1'}  \\
%\infer {\ifthen{\tt true}{e_2}{e_3} \stepo e_2}{\cdot} \and 
%\infer {\ifthen{\tt false}{e_2}{e_3} \stepo e_3}{\cdot}  \\
%\infer {\ifthen{e_1}{e_2}{e_3} \stept \ifthen{e_1'}{e_2}{e_3}}{e_1 \stept e_1' & e_2 \pct & e_3 \pct}  \\
%\infer {\ifthen{\tt true}{e_2}{e_3} \stept e_2}{e_2 \pct & e_3 \pct} \and 
%\infer {\ifthen{\tt false}{e_2}{e_3} \stept e_3}{e_2 \pct & e_3 \pct} \\
%\infer {\letin{x}{e_1}{e_2} \stepo \letin{x}{e_1'}{e_2}}{e_1 \stepo e_1'} \and
%\infer {\letin{x}{e_1}{e_2} \stepo [e_1/x]e_2}{e_1 \fco} \and
%\infer {\letin{x}{e_1}{e_2} \stept \letin{x}{e_1'}{e_2}}{e_1 \stept e_1'} \and
%\infer {\letin{x}{e_1}{e_2} \stept [e_1/x]e_2}{e_1 \fct} \and
%\infer {x \downtwo x}{\cdot} \and
%\infer {\gds (e_1,e_2) : A\times B}{\gds e_1 : A & \gds e_2 : B} \and
%\infer {\ifthen{e_1}{e_2}{e_3} \downone [v,r]}{e_1 \downone {\bf true} & e_2 \downone [v,r]} \and
%\infer {\ifthen{e_1}{e_2}{e_3} \downone [v,r]}{e_1 \downone {\bf false} & e_3 \downone [v,r]} \and
%\infer {\ifthen{e_1}{e_2}{e_3} \downtwo [v, \ifthen{r_1}{r_2}{r_3}]}{e_1 \downtwo [\valdum,r_1] & e_2 \downtwo [v_2,r_2] & e_3 \downtwo [v_3,r_3]} \and
%\infer {\next~e \downone [\valdum,r]}{e \downtwo r} \and
%\infer {\prev~e \downtwo r}{e \downone [\valdum, r]} \and
%\end{mathpar}
%\end{figure}

%\subsection{Good Theorems About the Language Alone}
%
%\begin{enumerate}
%\item Type Safety.
%\item Unidirectional Flow. [Essentially, you cannot write a program that goes back in time.]
%\item Erasure property.  [If you ignore the nexts and prevs, the result is weakly equivalent.] [This was proven for a denotational semantics in some Jones paper that I need to find...]
%\end{enumerate}

@@

\begin{figure*}
\caption{Term Splitting}
\label{fig:termSplit}
\begin{mathpar}
\infersplitone [\rmunit] 
	{\splitone{()}{\rmunit}{((),()),\_.()}} {\cdot} \and
\infersplittwo [\rmunit]
	{\splittwo{()}{\rmunit}{(),\_.()}} {\cdot} \and
\infersplitone [{\rm int}] 
	{\splitone{i}{\rm int}{(i,()),\_.()}} {\cdot} \and
\infersplittwo [{\rm int}]
	{\splittwo{i}{\rm int}{(),\_.i}} {\cdot} \and
\infersplitone [\times\mathrm{I}] 
	{\splitone {(e_1,e_2)}{A\times B}
		{ \begin{array}{l}
		\left(\letin{y_1}{c_1}{\letin{y_2}{c_2}{((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}\right), \\
		l.(\letin{l_1}{\pi_1 l}{r_1},\letin{l_2}{\pi_2 l}{r_2})
		\end{array}}}
	{ \splitonesub [1] A & \splitonesub [2] B} \and
\infersplittwo [\times\mathrm{I}] 
	{\splittwo {(e_1,e_2)}{A\times B}{(p_1,p_2), l.(\letin{l_1}{\pi_1~l}{r_1},\letin{l_2}{\pi_2~l}{r_2})}}
	{ \splittwosub [1] A & \splittwosub [2] B} \and
\infersplitone [\times\mathrm{E_1}] 
	{\splitone {\pi_1~e}{A}{(\pi_1(\pi_1c),\pi_2 c),l.\pi_1~r}}{\splitonesub{A\times B}} \and
\infersplittwo [\times\mathrm{E_1}] 
	{\splittwo{\pi_1~e}{A}{p,l.\pi_1~r} }{\splittwosub{A\times B}} \and
\infersplitone [\times\mathrm{E_2}] 
	{\splitone {\pi_2~e}{B}{(\pi_2(\pi_1c),\pi_2 c),l.\pi_2~r}}{\splitonesub{A\times B}} \and
\infersplittwo [\times\mathrm{E_2}] 
	{\splittwo{\pi_2~e}{B}{p,l.\pi_2~r} }{\splittwosub{A\times B}} \and
\infersplitone [\mathrm{unit}] 
	{\splitone {v}{A}{(x,()),\_.()}}{x:A^\bbone \in \Gamma} \and
\infersplittwo [\mathrm{unit}]
	 {\splittwo{v}{A}{(),\_.x}}{x:A^\bbtwo \in \Gamma} \\
\infersplitone [\mathrm{let}] 
	{\splitone{\letin{x}{e_1}{e_2}}{B}{\cdots}}
	{\splitonesub [1] A & \splitone [\Gamma,x:A^\bbone] {e_2}{B}{c_2,l_2.r_2}} \and
\infersplittwo [\mathrm{let}] 
	{\splittwo{\letin{x}{e_1}{e_2}}{B}{(p_1,p_2), l.\letin{x}{(\letin{l_1}{\pi_1 l}{r_1})}{\letin{l_2}{\pi_2 l}{r_2}}}}
	{\splittwosub [1] A & \splittwo [\Gamma,x:A^\bbtwo] {e_2}{B}{p_2,l_2.r_2}} \and
\infersplitone [\fut\mathrm{I}] 
	{\splitone{\next~e}{\fut A}{((),p),l.r}}{\splittwosub A} \and
\infersplittwo [\fut\mathrm{E}] 
	{\splittwo{\prev~e}{A} {\pi_2~c,l.r}}{\splitonesub {\fut A}} \and
\infersplitone [+\mathrm{E}]
	{\splitone {\left( \tallif {e_1}{e_2}{e_3} \right)}{A} 
	{
		\begin{array}{l}
		\left(\tallif {\pi_1c_1}
			{\letin{y}{c_2}{(\pi_1 y, \inl(\pi_2 y))}}
			{\letin{y}{c_3}{(\pi_1 y, \inr(\pi_2 y))}}
		\right), \\
		l.\left(\caseof{\pi_1 l}{l_2.r_2}{l_3.r_3}
		\right)
		\end{array}
	}}
	{\splitonesub [1] {\rm bool} & \splitonesub [2] A & \splitonesub [3] A } \and
\infersplittwo [+\mathrm{E}]
	{\splittwo {\left( \tallif {e_1}{e_2}{e_3}\right)}{A} 
	{(p_1,p_2,p_3), 
	l.\left(\tallif{(\letin {l_1}{\pi_1 l}{r_1})}{(\letin {l_2}{\pi_2 l}{r_2})} {(\letin {l_3}{\pi_3 l}{r_3})} \right)
	}}
	{\splittwosub [1] {\rm bool} & \splittwosub [2] A & \splittwosub [3] A } \and
\end{mathpar}
\end{figure*}

@@

The theorem we want to show has two parts, one for the $\bbone$ code, and one for the $\bbtwo$ code.  
\begin{center}
\begin{tabular}{l}
If $\splitone [\emptyC]{e}{A}{c,l.r}$ \\
and $\reduceone{e}{v}$ \\
and $c \Downarrow \valprod{t}{u}$ \\
and $[u/l]r \Downarrow w$ \\
then $v \vsplito [t,w]$
\end{tabular}
~~~
\begin{tabular}{l}
If $\splittwo[\emptyC]{e}{B}{p,l.r}$ \\
and $\reducetwo{e}{v}$ \\
and $p \Downarrow u$ \\
and $[u/l]r \Downarrow w$ \\
then $v \vsplits w$
\end{tabular}
\end{center}

@@

\begin{itemize}
\item {\bf Int.} Suppose that
\begin{enumerate}
\item $\splitone{i}{\rm int}{(i,()),\_.()}$
\end{enumerate}
Since reduction is a function, we also have
\begin{enumerate}
\item $\reduceone{i}{i}$
\item $\reduceone{(i,())}{\valprod {i}{()}}$
\item $\reduceone{[()/\_]()}{()}$
\end{enumerate}
And since value splitting is a function,
\begin{enumerate}
\item $i \vsplito [i,()]$
\end{enumerate}
Which is what we need to show.

>>>>

\begin{figure}
\caption{\corelang~Static Semantics}
\label{fig:coreStatics}
\begin{mathpar}
\infertypesone [\rmunit] 	{\ty {()}\rmunit}						{\cdot} \and
\infertypesone [int]			{\ty {i}{\rm int}}					{\cdot} \and
\infertypesone [bool]			{\ty {b} \rmbool}						{\cdot} \and
\infertypesone [hyp]			{\ty x A}								{x : A^\bbone \in \Gamma} \and
\infertypestwo [\rmunit]		{\ty {()}{\rmunit}}					{\cdot} \and
\infertypestwo [int]			{\ty {i}{\rm int}}					{\cdot} \and
\infertypestwo [bool]			{\ty {b} \rmbool}						{\cdot} \and
\infertypestwo [hyp]			{\ty x A}								{x : A^\bbtwo \in \Gamma} \and
\infertypesone [let]			{\ty {\letin{x}{e_1}{e_2}}{B}}		{\ty {e_1} A & \ty [\Gamma,x:A^\bbone] {e_2} B} \and
\infertypestwo [let]			{\ty {\letin{x}{e_1}{e_2}}{B}}		{\ty {e_1} A & \ty [\Gamma,x:A^\bbtwo] {e_2} B} \and
\infertypesone [\times I]	{\ty {(e_1,e_2)}{A\times B}}			{\ty {e_1} A & \ty {e_2} B} \and
\infertypesone [\times E_1]	{\ty {\pi_1~e} A}						{\ty e {A\times B}} \and
\infertypesone [\times E_2]	{\ty {\pi_2~e} B}						{\ty e {A\times B}} \and
\infertypestwo [\times I]	{\ty {(e_1,e_2)}{A\times B}}			{\ty {e_1} A & \typestwo {e_2} B} \and
\infertypestwo [\times E_1]	{\ty {\pi_1~e} A}						{\ty e {A\times B}} \and
\infertypestwo [\times E_2]	{\ty {\pi_2~e} B}						{\ty e {A\times B}} \and
\infertypesone [if] 			{\ty {\ifthen{e_1}{e_2}{e_3}} A}	{\ty {e_1} \rmbool & \ty {e_2} A & \ty {e_3} A} \and
\infertypestwo [if]			{\ty {\ifthen{e_1}{e_2}{e_3}} A}	{\ty {e_1} \rmbool & \ty {e_2} A & \ty {e_3} A} \and
\infertypesone [\fut I]		{\ty {\next~e}{\fut A}}				{\typestwo e A} \and
\infertypestwo [\fut E]		{\ty {\prev~e} A}						{\typesone e {\fut A}} \and
\end{mathpar}
\end{figure}

@@

\begin{figure}
\caption{\corelang~Erasure Dynamic Semantics}
\label{fig:erasureSemantics}
\begin{mathpar}
\inferreduceone [\rmunit]		{\red {()}{()}}							{\cdot} 											\and
\inferreduceone [int]				{\red {i}{i}}								{\cdot} 											\and
\inferreduceone [bool]			{\red {b}{b}}								{\cdot} 											\and
\inferreduceone [let]				{\red {\letin{x}{e_1}{e_2}}{v_2}}		{\sub [1] & \red{[v_1/x]e_2}{v_2}} 				\and
\inferreducetwo [\rmunit]		{\red {()}{()}}							{\cdot} 											\and
\inferreducetwo [int]				{\red {i}{i}}								{\cdot} 											\and
\inferreducetwo [bool]			{\red {b}{b}}								{\cdot} 											\and
\inferreducetwo [let]				{\red {\letin{x}{e_1}{e_2}}{v_2}}		{\sub [1] & \red{[v_1/x]e_2}{v_2}} 				\and
\inferreduceone [\times I]		{\red {(e_1,e_2)}{\valprod{v_1}{v_2}}}	{\sub [1] & \sub [2]} 							\and
\inferreduceone [\times E_1]		{\red {\pi_1~e}{v_1}}						{\red{e}{\valprod{v_1}{v_2}}} 					\and
\inferreduceone [\times E_2]		{\red {\pi_2~e}{v_2}}						{\red{e}{\valprod{v_1}{v_2}}} 					\and
\inferreducetwo [\times I]		{\red {(e_1,e_2)}{\valprod{v_1}{v_2}}}	{\sub [1] & \sub [2]}								\and
\inferreducetwo [\times E_1]		{\red {\pi_1~e}{v_1}}						{\red{e}{\valprod{v_1}{v_2}}} 					\and
\inferreducetwo [\times E_2]		{\red {\pi_2~e}{v_2}}						{\red{e}{\valprod{v_1}{v_2}}} 					\and
\inferreduceone [if_T] 			{\red {\ifthen{e_1}{e_2}{e_3}}{v}}		{\red{e_1}{\bf true}  & \red{e_2}{v}} 			\and
\inferreduceone [if_F] 			{\red {\ifthen{e_1}{e_2}{e_3}}{v}}		{\red{e_1}{\bf false} & \red{e_3}{v}}			\\
\inferreducetwo [if_T] 			{\red {\ifthen{e_1}{e_2}{e_3}}{v}}		{\red{e_1}{\bf true}  & \red{e_2}{v}} 			\and
\inferreducetwo [if_F] 			{\red {\ifthen{e_1}{e_2}{e_3}}{v}}		{\red{e_1}{\bf false} & \red{e_3}{v}}			\\
\inferreduceone [\fut I]			{\red {\next~e}{\valnext~v}}				{\reducetwosub}									\and
\inferreducetwo [\fut E]			{\red {\prev~e} v}						{\reduceone e {\valnext~v}} 
\end{mathpar}
\end{figure}

\begin{figure}
\caption{\corelang~Reference Semantics}
\label{fig:refSemantics}
\begin{mathpar}
\inferRedsplitone [\rmunit]		{\red {()}{()}{()}}												{\cdot} 												\and
\inferRedsplitone [int]			{\red {i}{i}{()}}													{\cdot} 												\and
\inferRedsplitone [bool]			{\red {b}{b}{()}}													{\cdot} 												\\
\inferSpeculation [\rmunit]		{\red {()}{()}}													{\cdot} 												\and
\inferSpeculation [int]			{\red {i}{i}}														{\cdot} 												\and
\inferSpeculation [bool]			{\red {b}{b}}														{\cdot} 												\\
\inferRedsplitone [let]			{\red {\letin{x}{e_1}{e_2}}{u_2}{\letin {x}{r_1}{r_2}}}		{\sub [1] & \red{[[u_1,x]/x]e_2}{u_2}{r_2}} 		\and
\inferSpeculation [let]			{\red {\letin{x}{e_1}{e_2}}{\letin{x}{r_1}{r_2}}}				{\sub [1] & \sub [2]}									\\
\inferRedsplitone [hyp]			{\red {[u,x]}{u}{x}}												{\cdot}													\and
\inferRedsplitone [\times I]		{\red {(e_1,e_2)}{\valprod{u_1}{u_2}}{\valprod{r_1}{r_2}}}	{\sub [1] & \sub [2]}									\and
\inferSpeculation [\times I]		{\red {(e_1,e_2)}{(r_1,r_2)}}									{\sub [1] & \sub [2]}									\and
\inferRedsplitone [\times E_1]	{\red {\pi_1~e}{u_1}{\pi_1 r}}									{\red {e}{\valprod{u_1}{u_2}}{r}}					\and
\inferSpeculation [\times E_1]	{\red {\pi_1~e}{\pi_1 r}}										{\red {e}{r}}											\and
\inferRedsplitone [\times E_2]	{\red {\pi_2~e}{u_2}{\pi_2 r}}									{\red {e}{\valprod{u_1}{u_2}}{r}} 					\and
\inferSpeculation [\times E_2]	{\red {\pi_2~e}{\pi_2 r}}										{\red {e}{r}} 											\and
\inferRedsplitone [if_T] 		{\red {\ifthen{e_1}{e_2}{e_3}}{u}{r_2}}							{\red{e_1}{\bf true }{r_1} & \red{e_2}{u}{r_2}}	\and
\inferRedsplitone [if_F] 		{\red {\ifthen{e_1}{e_2}{e_3}}{u}{r_3}}							{\red{e_1}{\bf false}{r_1} & \red{e_3}{u}{r_3}}	\\
\inferSpeculation [if] 			{\red {\ifthen{e_1}{e_2}{e_3}}{\ifthen{r_1}{r_2}{r_3}}}		{\sub [1] & \sub [2] & \sub [3]}					\\
\inferRedsplitone [\fut I]		{\red {\next~e}{()}{r}}											{\rstwosub}											\and
\inferSpeculation [\fut E]		{\red {\prev~e}{r}}												{\rsone {e}{()}{r}}
\end{mathpar}
\end{figure}

@@

\begin{figure*}
\caption{Term Splitting}
\label{fig:termSplit}
\newcommand {\projbind} [1] {\letin{l_{#1}}{\pi_{#1} l}{r_{#1}}}
\begin{mathpar}
\infersplitone [unit]			{\spl {()}{\rmunit}{((),()),\_.()}}															{\cdot} \and
\infersplittwo [unit]			{\spl {()}{\rmunit}{(),\_.()}}																{\cdot} \and
\infersplitone [int]			{\spl {i}{\rm int}{(i,()),\_.()}}															{\cdot} \and
\infersplittwo [int]			{\spl {i}{\rm int}{(),\_.i}}																	{\cdot} \and
\infersplitone [\times I] 	{\spl {(e_1,e_2)}{A\times B}
									{ \begin{array}{l}
									\left(
										\letin{y_1}{c_1}{\letin{y_2}{c_2}{
										((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}
									\right), \\
									l.(\projbind 1,\projbind 2)
									\end{array}}}																				{\sub 1 A & \sub 2 B} \and
\infersplittwo [\times I]	{\spl {(e_1,e_2)}{A\times B}{(p_1,p_2),l.(\projbind 1,\projbind 2)}}					{\sub 1 A & \sub 2 B} \and
\infersplitone [\times E_1]	{\spl {\pi_1~e}{A}{\letin{y}{c}{(\pi_1(\pi_1 y),\pi_2 y)},l.\pi_1~r}}					{\sub {} {A\times B}} \and
\infersplitone [\times E_2]	{\spl {\pi_2~e}{B}{\letin{y}{c}{(\pi_2(\pi_1 y),\pi_2 y)},l.\pi_2~r}}					{\sub {} {A\times B}} \and
\infersplittwo [\times E_1]	{\spl {\pi_1~e}{A}{p,l.\pi_1~r}}															{\sub {} {A\times B}} \and
\infersplittwo [\times E_2]	{\spl {\pi_2~e}{B}{p,l.\pi_2~r}}															{\sub {} {A\times B}} \and
\infersplitone [hyp]			{\spl {v}{A}{(x,()),\_.()}}																	{x:A^\bbone \in \Gamma} \and
\infersplittwo [hyp]			{\spl {v}{A}{(),\_.x}}																		{x:A^\bbtwo \in \Gamma} \\
\infersplitone [let]			{\spl {\letin{x}{e_1}{e_2}}{B}{\cdots}}														{\sub 1 A & \sub [\Gamma,x:A^\bbone] 2 B} \and
\infersplittwo [let]			{\spl {\letin{x}{e_1}{e_2}}{B}{(p_1,p_2), l.\letin{x}{(\projbind 1)}{\projbind 2}}}	{\sub 1 A & \sub [\Gamma,x:A^\bbtwo] 2 B} \and
\infersplitone [\fut I]		{\spl {\next~e}{\fut A}{((),p),l.r}}														{\splittwosub {} A} \and
\infersplittwo [\fut E]		{\spl {\prev~e}{A} {\pi_2~c,l.r}}															{\splitonesub {} {\fut A}} \and
\infersplitone [if]			{\spl {\left( \tallif {e_1}{e_2}{e_3} \right)}{A} 
								{\begin{array}{l}
									\left(\tallif {\pi_1 c_1}
										{\letin{y}{c_2}{(\pi_1 y, \inl(\pi_2 y))}}
										{\letin{y}{c_3}{(\pi_1 y, \inr(\pi_2 y))}}
									\right), \\
									l.(\caseof{\pi_1 l}{l_2.r_2}{l_3.r_3})
								\end{array}}}																					{\sub 1 \rmbool & \sub 2 A & \sub 3 A } \and
\infersplittwo [if]			{\spl {\left(\tallif {e_1}{e_2}{e_3}\right)}{A} 
								{(p_1,p_2,p_3),l.\left(\tallif{(\projbind 1)}{(\projbind 2)}{(\projbind 3)}\right)}}	{\sub 1 \rmbool & \sub 2 A & \sub 3 A } \and
\end{mathpar}
\end{figure*}

@@

The theorem we want to show has two parts, one for the $\bbone$ code, and one for the $\bbtwo$ code.  
\begin{center}
\begin{tabular}{l}
If $\splitone [\emptyC]{e}{A}{c,l.r}$ \\
and $\rsone{e}{v}{q}$ \\
and $\reducetwo{q}{w}$ \\
then $\reduce{c}{\valprod{v}{u}}$ \\
and $\reduce{[u/l]r}{w}$
\end{tabular}
~~~
\begin{tabular}{l}
If $\splittwo[\emptyC]{e}{B}{p,l.r}$ \\
and $\rstwo{e}{q}$ \\
and $\reducetwo{q}{w}$ \\
then $\reduce{p}{u}$ \\
and $\reduce{[u/l]r}{w}$ 
\end{tabular}
\end{center}

@@

\begin{itemize}
\item {\bf Int.} Suppose that
\begin{enumerate}
\item $\splitone{i}{\rm int}{(i,()),\_.()}$
\item $\rsone{i}{v}{q}$
\item $\reducetwo{r}{w}$
\end{enumerate}
By inversion of reduction relations, we know
\begin{enumerate}
\item $v$ has the form $i$
\item $q$ has the form $()$
\item $w$ has the form $()$
\end{enumerate}
It is directly provable that,
\begin{enumerate}
\item $\reduce{(i,())}{\valprod{i}{()}}$
\item $\reduce{[()/l]()}{()}$
\end{enumerate}
Which is what we need to show.

\item {\bf Tuple.} Suppose that
\begin{enumerate}
\item $\splitone {(e_1,e_2)}{A\times B}
		{ \begin{array}{l}
		(\letin{y_1}{c_1}{\letin{y_2}{c_2}{((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}), \\
		l.(\letin{l_1}{\pi_1 l}{r_1},\letin{l_2}{\pi_2 l}{r_2})
		\end{array}} $
\item $\rsone{(e_1,e_2)}{v}{q}$
\item $\reducetwo{q}{w}$
\end{enumerate}
By inversion of the splitting and reduction rules, we know
\begin{enumerate}
\item $\splitone{e_1}{A}{c_1,l_1.r_1}$ and $\splitone{e_2}{B}{c_2,l_2.r_2}$
\item $v$ has the form $\valprod{v_1}{v_2}$
\item $q$ has the form $(q_1,q_2)$
\item $\rsone{e_1}{v_1}{q_1}$ and $\rsone{e_2}{v_2}{q_2}$
\item $w$ has the form $\valprod{w_1}{w_2}$
\item $\reducetwo{q_1}{w_1}$ and $\reducetwo{q_2}{w_2}$
\end{enumerate}
This gives us enough to invoke the inductive hypothesis for the each side, yielding
\begin{enumerate}
\item $\reduce{c_1}{\valprod{v_1}{u_1}}$ 
\item $\reduce{[u_1/l_1]r_1}{w_1}$ 
\item $\reduce{c_2}{\valprod{v_2}{u_2}}$ 
\item $\reduce{[u_2/l_2]r_2}{w_2}$ 
\end{enumerate}
It is now directly provable from the evaluation rules that
\begin{enumerate}
\item $\reduce{\letin{y_1}{c_1}{\letin{y_2}{c_2}{((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}}{\valprod {\valprod{v_1}{v_2}}{\valprod{u_1}{u_2}}}$
\item $\reduce{[\valprod{u_1}{u_2}/l](\letin{l_1}{\pi_1 l}{r_1},\letin{l_2}{\pi_2 l}{r_2})}{\valprod{w_1}{w_2}}$
\end{enumerate}
Which is what we wanted to show.

\item {\bf Projection.} Suppose that
\begin{enumerate}
\item $\splitone {\pi_1~e}{A}{\letin{y}{c}{(\pi_1(\pi_1 y),\pi_2 y)},l.\pi_1~r}$
\item $\rsone{\pi_1~e}{v}{q}$
\item $\reducetwo{q}{w}$
\end{enumerate}
By inversion of the splitting and reduction rules, we have
\begin{enumerate}
\item $\splitone {e}{A \times B}{c,l.r}$
\item $q$ has the form $\pi_1~q'$
\item $\rsone{e}{\valprod{v}{v_2}}{q'}$
\item $\reducetwo{q'}{\valprod{w}{w_2}}$
\end{enumerate}
This is enough to invoke our inductive hypothesis, yielding
\begin{enumerate}
\item $\reduce{c}{\valprod{\valprod{v}{v_2}}{u}}$
\item $\reduce{[u/l]r}{\valprod{w}{w_2}}$
\end{enumerate}
From this, it can be directly proven that
\begin{enumerate}
\item $\reduce{\letin{y}{c}{(\pi_1(\pi_1 y),\pi_2 y)}}{\valprod{v}{u}}$
\item $\reduce{[u/l](\pi_1~r)}{w}$
\end{enumerate}
Which is what we wanted to show.
